Matching Bullets

### 2015-12-15 imaging-paper.Rnw
acknowledgement

The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks. This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system. A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2014} came to the same conclusions based on the EvoFinder system. 
The NRC report also found, that results from 2d images are improved upon by matches based on 3d images. This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.

@@

This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).

>>>>

The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks. This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system. A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2015} came to the same conclusions based on the EvoFinder system. 
The NRC report also found, that results from 2d images are improved upon by matches based on 3d images. This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.

@@

This research was partially funded through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
